Imputation of Incomplete Multilevel Data with \pkg{mice}

### 2022-04-07 Manuscript/Imputation_of_Incomplete_Multilevel_Data.Rmd
Add forest plot effect ct

We illustrate how to impute incomplete multilevel data by means of a case study: `impact` from the \pkg{metamisc} package [empirical data on traumatic brain injuries, $n = 11,022$ patients across $N = 15$ studies, @metamisc]. `impact` is traumatic brain injury data with patients, $n = 11022$, clustered in studies, $N = 15$. With the following 11 variables:

@@

[TODO: add more info about the complete data.] The analysis model for this dataset is a prediction model with `mort` as the outcome. In this tutorial we'll estimate the adjusted prognostic effect of `ct` on unfortunate outcomes. The estimand is the adjusted odds ratio for `ct`, after including `type`, `age` `motor_score` and `pupil` into the analysis model: 
```{r mod}
mod <- mort ~ 1 + type + age + motor_score + pupil + ct + (1 | name)
```

@@

Load the complete and incomplete data into the R workspace:
```{r data, eval = FALSE}
data("impact", package = "metamisc")
dat <- read.table("link/to/the/data.txt")
```

@@

We will use the following estimates as comparative truth in this tutorial [TODO: make this a table or forest plot instead, to see if there is heterogeneity in the association of `ct` with `mort`]:
```{r truth}
fit <- glmer(mod, family = "binomial", data = impact) 
tidy(fit, conf.int = TRUE, exponentiate = TRUE) 
```

@@

To explore the missingness, it is wise to look at the missing data pattern:
```{r pattern, fig.height=7.1}
plot_pattern(dat, rotate = TRUE)
```

@@

We should impute the variables `ct` and `pupil` and any auxiliary variables we might want to use to impute these incomplete analysis model variables. We can evaluate which variables may be useful auxiliaries by plotting the pairwise complete correlations:
```{r impact_corr}
plot_corr(dat, rotate = TRUE)
```

This shows us that `hypox` and `hypot` would not be useful auxiliary variables for imputing `ct`. Depending on the minimum required correlation, `tsah` could be useful, while `edh` has the strongest correlation with `ct` out of all the variables in the data and should definitely be included in the imputation model. For the imputation of `pupil`, none of the potential auxiliary variables has a very strong relation, but `hypots` could be used. We conclude that we can exclude `hypox` from the data, since this is neither an analysis model variable nor an auxiliary variable for imputation:
```{r}
dat <- select(dat, !hypox)
```

@@

As previously stated, complete case analysis lowers statistical power and may bias results. The complete case analysis estimates are:
```{r cca}
fit <- glmer(mod, family = "binomial", data = na.omit(dat)) 
tidy(fit, conf.int = TRUE, exponentiate = TRUE) 
```

>>>>

We illustrate how to impute incomplete multilevel data by means of a case study: `impact` from the \pkg{metamisc} package [empirical data on traumatic brain injuries, $n = 11,022$ units across $N = 15$ clusters, @metamisc]. [TODO: add more info about the complete data.] The `impact` data set contains traumatic brain injury data on $n = 11022$ patients clustered in $N = 15$ studies with the following 11 variables:

@@

The analysis model for this dataset is a prediction model with `mort` as the outcome. In this tutorial we'll estimate the adjusted prognostic effect of `ct` on unfortunate outcomes. The estimand is the adjusted odds ratio for `ct`, after including `type`, `age` `motor_score` and `pupil` into the analysis model: 
```{r mod}
mod <- mort ~ 1 + type + age + motor_score + pupil + ct + (1 | name) # TODO: add random effects ct?
```

@@

Load the complete and incomplete data into the R workspace:
```{r data, eval = FALSE}
data("impact", package = "metamisc")      # complete data
dat <- read.table("link/to/the/data.txt") # incomplete data
```

@@

We will use the following estimates as comparative truth in this tutorial [TODO: make this a table or forest plot instead, to see if there is heterogeneity in the association of `ct` with `mort`]:
```{r truth}
fit <- glmer(mod, family = "binomial", data = impact) # fit the model
tidy(fit, conf.int = TRUE, exponentiate = TRUE)       # print estimates
```

```{r forest, echo=FALSE}
fits <-
  purrr::map_dfr(
    unique(impact$name),
    ~ {
      glm(
        mort ~ 1 + age + motor_score + pupil + ct,
        family = "binomial",
        data = filter(impact, name == .x)
      )
    } %>%
      broom::tidy(conf.int = TRUE, exponentiate = TRUE) %>%
      .[, c("term", "estimate", "conf.low", "conf.high")] %>%
      cbind(name = as.character(.x), .)
  )
filter(fits, term %in% c("ctIII")) %>% #, "ctIV/V"
  ggplot(aes(y = name, x = estimate)) +
  geom_vline(xintercept = 1, size = 1.5, color = "grey95") +
  geom_point() +
  geom_linerange(aes(y = name, xmin = conf.low, xmax = conf.high)) +
  labs(y = "Study", x = "Odds ratio", color = NULL) +
  theme_classic()
```

@@

To explore the missingness, it is wise to look at the missing data pattern:
```{r pattern, fig.height=7.1}
plot_pattern(dat, rotate = TRUE)  # plot missingness pattern
```

@@

We should impute the variables `ct` and `pupil` and any auxiliary variables we might want to use to impute these incomplete analysis model variables. We can evaluate which variables may be useful auxiliaries by plotting the pairwise complete correlations:
```{r impact_corr}
plot_corr(dat, rotate = TRUE) # plot correlations 
```

This shows us that `hypox` and `hypot` would not be useful auxiliary variables for imputing `ct`. Depending on the minimum required correlation, `tsah` could be useful, while `edh` has the strongest correlation with `ct` out of all the variables in the data and should definitely be included in the imputation model. For the imputation of `pupil`, none of the potential auxiliary variables has a very strong relation, but `hypots` could be used. We conclude that we can exclude `hypox` from the data, since this is neither an analysis model variable nor an auxiliary variable for imputation:
```{r}
dat <- select(dat, !hypox)  # remove variable
```

@@

As previously stated, complete case analysis lowers statistical power and may bias results. The complete case analysis estimates are:
```{r cca}
fit <- glmer(mod, family = "binomial", data = na.omit(dat)) # fit the model
tidy(fit, conf.int = TRUE, exponentiate = TRUE)             # print estimates
```

### 2022-04-07 Manuscript/Imputation_of_Incomplete_Multilevel_Data.tex
Add forest plot effect ct

We illustrate how to impute incomplete multilevel data by means of a
case study: \texttt{impact} from the \pkg{metamisc} package
\citep[empirical data on traumatic brain injuries, \(n = 11,022\)
patients across \(N = 15\) studies,][]{metamisc}. \texttt{impact} is
traumatic brain injury data with patients, \(n = 11022\), clustered in
studies, \(N = 15\). With the following 11 variables:

@@

{[}TODO: add more info about the complete data.{]} The analysis model
for this dataset is a prediction model with \texttt{mort} as the
outcome. In this tutorial we'll estimate the adjusted prognostic effect
of \texttt{ct} on unfortunate outcomes. The estimand is the adjusted
odds ratio for \texttt{ct}, after including \texttt{type}, \texttt{age}
\texttt{motor\_score} and \texttt{pupil} into the analysis model:

@@

\begin{CodeChunk}
\begin{CodeInput}
R> data("impact", package = "metamisc")
R> dat <- read.table("link/to/the/data.txt")
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> fit <- glmer(mod, family = "binomial", data = impact) 
R> tidy(fit, conf.int = TRUE, exponentiate = TRUE) 
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 11 x 9
   effect group term  estimate std.error statistic    p.value conf.low conf.high
   <chr>  <chr> <chr>    <dbl>     <dbl>     <dbl>      <dbl>    <dbl>     <dbl>
 1 fixed  <NA>  (Int~    0.101   0.0173     -13.4   4.94e- 41   0.0723     0.141
 2 fixed  <NA>  type~    0.713   0.123       -1.96  5.01e-  2   0.509      1.00 
 3 fixed  <NA>  age      1.03    0.00165     20.2   2.13e- 90   1.03       1.04 
 4 fixed  <NA>  moto~    0.553   0.0386      -8.50  1.95e- 17   0.482      0.634
 5 fixed  <NA>  moto~    0.405   0.0289     -12.7   8.43e- 37   0.352      0.466
 6 fixed  <NA>  moto~    0.275   0.0202     -17.6   1.67e- 69   0.239      0.318
 7 fixed  <NA>  pupi~    3.73    0.230       21.4   2.09e-101   3.31       4.21 
 8 fixed  <NA>  pupi~    1.87    0.133        8.80  1.36e- 18   1.63       2.15 
 9 fixed  <NA>  ctIII    2.25    0.157       11.6   5.12e- 31   1.96       2.58 
10 fixed  <NA>  ctIV~    2.30    0.136       14.0   9.47e- 45   2.05       2.58 
11 ran_p~ name  sd__~    0.277  NA           NA    NA          NA         NA    
\end{CodeOutput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plot_pattern(dat, rotate = TRUE)
\end{CodeInput}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plot_corr(dat, rotate = TRUE)
\end{CodeInput}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> dat <- select(dat, !hypox)
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> fit <- glmer(mod, family = "binomial", data = na.omit(dat)) 
R> tidy(fit, conf.int = TRUE, exponentiate = TRUE) 
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 11 x 9
   effect  group term  estimate std.error statistic   p.value conf.low conf.high
   <chr>   <chr> <chr>    <dbl>     <dbl>     <dbl>     <dbl>    <dbl>     <dbl>
 1 fixed   <NA>  (Int~   0.0863   0.0182     -11.6   2.99e-31   0.0571     0.130
 2 fixed   <NA>  type~   0.757    0.137       -1.54  1.22e- 1   0.531      1.08 
 3 fixed   <NA>  age     1.03     0.00265     12.9   7.40e-38   1.03       1.04 
 4 fixed   <NA>  moto~   0.651    0.0732      -3.82  1.34e- 4   0.522      0.811
 5 fixed   <NA>  moto~   0.489    0.0555      -6.30  2.97e-10   0.391      0.611
 6 fixed   <NA>  moto~   0.274    0.0321     -11.0   2.28e-28   0.218      0.345
 7 fixed   <NA>  pupi~   3.20     0.317       11.7   8.18e-32   2.63       3.88 
 8 fixed   <NA>  pupi~   1.75     0.195        5.06  4.27e- 7   1.41       2.18 
 9 fixed   <NA>  ctIII   2.41     0.268        7.89  3.05e-15   1.94       2.99 
10 fixed   <NA>  ctIV~   2.30     0.214        8.95  3.55e-19   1.92       2.76 
11 ran_pa~ name  sd__~   0.230   NA           NA    NA         NA         NA    
\end{CodeOutput}
\end{CodeChunk}

>>>>

We illustrate how to impute incomplete multilevel data by means of a
case study: \texttt{impact} from the \pkg{metamisc} package
\citep[empirical data on traumatic brain injuries, \(n = 11,022\) units
across \(N = 15\) clusters,][]{metamisc}. {[}TODO: add more info about
the complete data.{]} The \texttt{impact} data set contains traumatic
brain injury data on \(n = 11022\) patients clustered in \(N = 15\)
studies with the following 11 variables:

@@

The analysis model for this dataset is a prediction model with
\texttt{mort} as the outcome. In this tutorial we'll estimate the
adjusted prognostic effect of \texttt{ct} on unfortunate outcomes. The
estimand is the adjusted odds ratio for \texttt{ct}, after including
\texttt{type}, \texttt{age} \texttt{motor\_score} and \texttt{pupil}
into the analysis model:

@@

\begin{CodeChunk}
\begin{CodeInput}
R> data("impact", package = "metamisc")      # complete data
R> dat <- read.table("link/to/the/data.txt") # incomplete data
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> fit <- glmer(mod, family = "binomial", data = impact) # fit the model
R> tidy(fit, conf.int = TRUE, exponentiate = TRUE)       # print estimates
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 11 x 9
   effect group term  estimate std.error statistic    p.value conf.low conf.high
   <chr>  <chr> <chr>    <dbl>     <dbl>     <dbl>      <dbl>    <dbl>     <dbl>
 1 fixed  <NA>  (Int~    0.101   0.0173     -13.4   4.94e- 41   0.0723     0.141
 2 fixed  <NA>  type~    0.713   0.123       -1.96  5.01e-  2   0.509      1.00 
 3 fixed  <NA>  age      1.03    0.00165     20.2   2.13e- 90   1.03       1.04 
 4 fixed  <NA>  moto~    0.553   0.0386      -8.50  1.95e- 17   0.482      0.634
 5 fixed  <NA>  moto~    0.405   0.0289     -12.7   8.43e- 37   0.352      0.466
 6 fixed  <NA>  moto~    0.275   0.0202     -17.6   1.67e- 69   0.239      0.318
 7 fixed  <NA>  pupi~    3.73    0.230       21.4   2.09e-101   3.31       4.21 
 8 fixed  <NA>  pupi~    1.87    0.133        8.80  1.36e- 18   1.63       2.15 
 9 fixed  <NA>  ctIII    2.25    0.157       11.6   5.12e- 31   1.96       2.58 
10 fixed  <NA>  ctIV~    2.30    0.136       14.0   9.47e- 45   2.05       2.58 
11 ran_p~ name  sd__~    0.277  NA           NA    NA          NA         NA    
\end{CodeOutput}
\end{CodeChunk}

\begin{CodeChunk}

\begin{center}\includegraphics{Imputation_of_Incomplete_Multilevel_Data_files/figure-latex/forest-1} \end{center}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plot_pattern(dat, rotate = TRUE)  # plot missingness pattern
\end{CodeInput}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> plot_corr(dat, rotate = TRUE) # plot correlations 
\end{CodeInput}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> dat <- select(dat, !hypox)  # remove variable
\end{CodeInput}
\end{CodeChunk}

@@

\begin{CodeChunk}
\begin{CodeInput}
R> fit <- glmer(mod, family = "binomial", data = na.omit(dat)) # fit the model
R> tidy(fit, conf.int = TRUE, exponentiate = TRUE)             # print estimates
\end{CodeInput}
\begin{CodeOutput}
# A tibble: 11 x 9
   effect  group term  estimate std.error statistic   p.value conf.low conf.high
   <chr>   <chr> <chr>    <dbl>     <dbl>     <dbl>     <dbl>    <dbl>     <dbl>
 1 fixed   <NA>  (Int~   0.0863   0.0182     -11.6   2.99e-31   0.0571     0.130
 2 fixed   <NA>  type~   0.757    0.137       -1.54  1.22e- 1   0.531      1.08 
 3 fixed   <NA>  age     1.03     0.00265     12.9   7.40e-38   1.03       1.04 
 4 fixed   <NA>  moto~   0.651    0.0732      -3.82  1.34e- 4   0.522      0.811
 5 fixed   <NA>  moto~   0.489    0.0555      -6.30  2.97e-10   0.391      0.611
 6 fixed   <NA>  moto~   0.274    0.0321     -11.0   2.28e-28   0.218      0.345
 7 fixed   <NA>  pupi~   3.20     0.317       11.7   8.18e-32   2.63       3.88 
 8 fixed   <NA>  pupi~   1.75     0.195        5.06  4.27e- 7   1.41       2.18 
 9 fixed   <NA>  ctIII   2.41     0.268        7.89  3.05e-15   1.94       2.99 
10 fixed   <NA>  ctIV~   2.30     0.214        8.95  3.55e-19   1.92       2.76 
11 ran_pa~ name  sd__~   0.230   NA           NA    NA         NA         NA    
\end{CodeOutput}
\end{CodeChunk}
